Bootstrapping Multi-atlas Segmentation Using Multiple Automatically Generated Templates for the Segmentation of the Whole Hippocampus and Subfields

### 2013-12-20 paper.Rnw
Added subregion volume p-values

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
This experiment explores \mb{} segmentation of hippocampal subfields. To achieve
this, a leave-one-out validation was conducted in which lower-resolution images
($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
remaining high-resolution Winterburn atlas subjects' images.

In general, across hippocampal subfields the percent error in volume of \mb{}
segmentations (relative to the full-resolution Winterburn atlas segmentation)
compares favourably to the error resulting from image resampling (Figure
\ref{fig:WAval-vol-boxplot}). In particular, the CA1, CA4/DG, and SR/SL/SM
subfields all show a percent error in volume that is at or lower than
resampling error. The Subiculum and CA2/CA3 subfields show distinctly larger
percent error in volume than is found through resampling.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: compare resampling
\begin{figure}[h!]
<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
data = subset(read.csv('data/WAval.csv'))
casted = dcast(melt(data, id.vars=c("file","version")), file + variable ~ version)
casted = rename(casted, c("variable"="region"))

levels(error$region) <-list("CA1"="X1", "CA1"="X101", 
                            "Subiculum"="X2","Subiculum"="X102", 
                            "CA4/DG"="X4","CA4/DG"="X104", 
                            "CA2/CA3"="X5","CA2/CA3"="X105", 
                            "SR/SL/SM"="X6","SR/SL/SM"="X106")
melted = melt(error,id.vars=c("file", "region"))
melted = rename(melted, c("variable"="measure", "value"="percenterr"))
levels(melted$measure) <- list("Subsampled"="resampled",
                               "MAGeT on Subsampled" = "mb_0.9mm",
                               "MAGeT on T1 BRAVO"="mb_bravo")

ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
  geom_hline(aes(yintercept=0), colour="grey", size=1) + 
  geom_boxplot() +
  labs(y="Percent error in volume", x="Subfield") +
  scale_colour_discrete("") + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Average percent error in hemispheric subfield volume.} 
  Percent error is measured against the volumes of the unmodified Winterburn
  atlas subfield segmentations.
  {\bf Subsampled} are volumes of the manual segmentations of the Winterburn
  atlases after resampling to $0.9mm^3$ voxels.  
  {\bf MAGeT on Subsampled} volumes are \mb{} segmentations of the Winterburn atlas
  images after resampling to $0.9mm^3$ voxels.
  {\bf MAGeT on T1 BRAVO} volumes are \mb{} segmentations of T1 BRAVO images
  ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
  subjects.} 
  \label{fig:WAval-vol-boxplot}
\end{figure}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
<<WAval-prep, cache=T>>=
data = subset(read.csv('data/WAval.csv'))
casted = dcast(melt(data, id.vars=c("file","version")), file + variable ~ version)
casted = rename(casted, c("variable"="region"))

levels(error$region) <-list("CA1"="X1", "CA1"="X101", 
                            "Subiculum"="X2","Subiculum"="X102", 
                            "CA4/DG"="X4","CA4/DG"="X104", 
                            "CA2/CA3"="X5","CA2/CA3"="X105", 
                            "SR/SL/SM"="X6","SR/SL/SM"="X106")
melted = melt(error,id.vars=c("file", "region"))
melted = rename(melted, c("variable"="measure", "value"="percenterr"))
levels(melted$measure) <- list("Subsampled atlas labels"="resampled",
                               "MAGeT on subsampled atlas images" = "mb_0.9mm",
                               "MAGeT on T1 BRAVO images"="mb_bravo")

error.t.tests = ddply(error, c("region"), function (df) { 
  data.frame( 
    bravo = round(t.test(df$resampled, df$mb_bravo, paired=T)$p.value,3), 
    atlas = round(t.test(df$resampled, df$mb_0.9mm, paired=T)$p.value,3)
  )}) 
@
This experiment explores \mb{} segmentation of hippocampal subfields. To achieve
this, a leave-one-out validation was conducted in which lower-resolution images
($0.9mm^3$ voxels) of each Winterburn atlas subject were segmented using the
remaining high-resolution Winterburn atlas subjects' images.

In general, across hippocampal subfields the percent error in volume of \mb{}
segmentations (relative to the full-resolution Winterburn atlas segmentation)
compares favourably to the error resulting from image resampling (Figure
\ref{fig:WAval-vol-boxplot}). In particular, there is no significant difference 
between percent error of \mb{} volumes compared to image resampling error for the 
CA1, Subiculum, and CA2/CA3 subregions. The \mb{} volumes of CA4/DG and SR/SL/SM 
subregions show both significantly different (lower) percent error (p < 0.001).

%the CA1, CA4/DG, and SR/SL/SM
%subfields all show a percent error in volume that is at or lower than
%resampling error. The Subiculum and CA2/CA3 subfields show distinctly larger
%percent error in volume than is found through resampling.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: compare resampling
\begin{figure}[h!]
<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
ggplot(melted, aes(y=percenterr,x=measure,colour=measure)) +
  facet_grid(.~region) + 
  geom_hline(aes(yintercept=0), colour="grey", size=1) + 
  geom_boxplot() + 
  geom_point(group=1, position="dodge") + 
  labs(y="Unilateral subregion volume percent error", x="Subfield") +
  theme(axis.ticks = element_blank(), axis.text.x = element_blank()) + 
  scale_colour_discrete("") + 
  theme(legend.direction = "vertical", legend.position = "bottom")
@
  \caption{{\bf Percent error in unilateral subfield volume.} 
  Percent error is measured against the volumes of the unmodified Winterburn
  atlas subfield segmentations.
  {\bf Subsampled atlas labels} volumes of the manual segmentations of the Winterburn
  atlases after resampling to $0.9mm^3$ voxels.  
  {\bf MAGeT on subsampled atlas images} volumes are \mb{} segmentations of the Winterburn atlas
  images after resampling to $0.9mm^3$ voxels.
  {\bf MAGeT on T1 BRAVO} volumes are \mb{} segmentations of T1 BRAVO images
  ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
  subjects.} 
  \label{fig:WAval-vol-boxplot}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 1yr Complete Dataset Package Totals & Failures
%                        
<<ADNI-xvals-pvalue, echo=F, results="asis",cache=FALSE>>=
latex(error.t.tests, file="", size="scriptsize",landscape=FALSE, ctable=TRUE,
    caption="\\textbf{t-test p-values whem comparing to resampled percent error.}", 
    caption.loc = c("top"), rowname=NULL, 
    label="tab:resampled-percent-error")
@
